---
title: Par치bolas con ggplot
author: Rafa
date: '2018-08-19'
slug: parabolas-con-ggplot
categories: []
tags: []
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE)
library(tidyverse)
library(latex2exp)
my_theme <- theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))
theme_set(my_theme)
```
## Gr치fico de un Par치bola en ggplot

Hay dos formas de graficar una funci칩n en `ggplot`: usando `stat_function` y usando `annotate`.
```{r, warning = FALSE}
t <- seq(-5, 5, length.out = 100)
g <- ggplot() + 
  annotate("path", 
           x=  t,
           y =  t^2) + 
  xlim(-5, 5) +
  geom_segment(aes(x=-5, xend = 5, y=0, yend=0),
               size = 0.5,
               arrow = arrow(length = unit(0.2, "cm"))) + # Eje X
  geom_segment(aes(x=0, xend=0, y=-1, yend=25),
               size = 0.5,
               lineend = "butt",
               arrow = arrow(length = unit(0.2, "cm"))) + # Eje Y 
  ggtitle(TeX('$y=x^2$')) + 
  theme(
    plot.title = element_text(hjust = 0.5)
  ) +
  NULL
g
```


Or stat_fun:
  
  
```{r}
library(latex2exp)

fun.1 <- function(x) { x^2 }
p <- ggplot(data_frame(x=0), aes(x=x))
p + 
  stat_function(fun = fun.1) + xlim(-5, 5) + 
  geom_segment(aes(x=-5, xend = 5, y=0, yend=0),
               size = 0.5,
               arrow = arrow(length = unit(0.2, "cm"))) + # Eje X
  geom_segment(aes(x=0, xend=0, y=-1, yend=25),
               size = 0.5,
               lineend = "butt",
               arrow = arrow(length = unit(0.2, "cm"))) + # Eje Y 
  ggtitle(TeX('$y=x^2$')) + 
  theme(
    plot.title = element_text(hjust = 0.5)
  ) +
  NULL
```

The set of point equidistante from a point (focus) and a line (directrix).

### Equation

$(x-h)^2=4p(y-k), p \neq 0$
  With vertex at $(h,k)$ vertical axis and directrix $y=k-p$.

### Example
$x^2=16y$ is a parabola with vertex at $(0,0)$, $p = 4$ and focus in $(0,4)$.

```{r}
focus <- list(x=0, y=4, lab = "Focus")
vertex <- list(x=0, y=0, lab = "Vertex")
notable_points <- bind_rows(focus, 
                            vertex)

pt <- data_frame(x=5, y=round(5^2/16, 2))

p <- ggplot(data_frame(x=0), aes(x=x)) +
  geom_segment(aes(x=-9, xend = 9, y=0, yend=0),
               size = 0.5,
               arrow = arrow(length = unit(0.2, "cm"))) + # Eje X
  geom_segment(aes(x=0, xend=0, y=-5, yend=10),
               size = 0.5,
               lineend = "butt",
               arrow = arrow(length = unit(0.2, "cm"))) + # Eje Y
  stat_function(fun = function(x) { x^2 / 16}) + 
  xlim(-9, 9) + 
  ylim(-5, 10)

p_annotated <- p +
  geom_point(data = notable_points, aes(x=x, y=y)) + 
  geom_text(data=notable_points, 
            size = 3,
            nudge_y = 1.25,
            nudge_x = 1.25,
            aes(label = paste0(lab, "\n(", x, ",", y, ")"), 
                x = x, y = y)) + 
  geom_hline(yintercept = -4, linetype = "dashed") + 
  annotate("text", x= 3, y= -3.3, label = "Directrix", size = 3) +
  
  ggtitle(TeX('$y=x^2/16$')) + 
  theme(
    plot.title = element_text(hjust = 0.5)
  ) +
  NULL

x <- 5
p_equipoints <- p_annotated + geom_point(data = pt, aes(x=x, y=y)) + 
  geom_segment(aes(x= focus$x, y=focus$y, xend=pt$x, yend=pt$y),
               linetype = "dashed") + 
  geom_segment(aes(x=pt$x, y=pt$y, xend=pt$x, yend = -4),
               linetype = "dashed") + 
  annotate("text", 
           label = "phantom(0) == phantom(0)", 
           parse = TRUE,
           angle = 60,
           size = 5,
           x = 2.5, y= 2.85) + 
  annotate("text", 
           label = "phantom(0) == phantom(0)", 
           parse = TRUE,
           size = 5,
           x = 5, y= -2) + 
  annotate("text")
p_equipoints
```




## Another Example
Complete the
\begin{align}
y^2+6y+8x+25 &= 0\\
y^2+6y+9+8x+25-9  &= 0\\
(y+3)^2 &= -16-8x\\
(y+3)^2 &= 4(-2)(x+2)
\end{align}

Vertex is $(-2, -3)$, p is $-2$, focus is $(-4,-3)$ directrix is $x=0$. But this parabola is not a function. How do we plot it?
  Break it in two pieces.
\begin{align}
(y+3)^2 &= (-8)(x+2)\\
y     &= -3 \pm \sqrt{-8(x+2)}\\
y_1  &=-3 + \sqrt{-8(x+2)}\\
y_2 &=-3 - \sqrt{-8(x+2)}\\
\end{align}

Graph those $y1$ and $y2$:
  
```{r}
x <- seq(-5, 0, length.out=1e3)
lbs <- list(bquote(-3 + sqrt(-8*(x+2))), # investigar bquote!
            bquote(-3 - sqrt(-8*(x+2))))

g <- ggplot(data_frame(x=x), 
            aes(x=x, color = color)) + 
  stat_function(data = data_frame(x=x, color = factor(1)),
                fun = function(x) { -3 + sqrt(-8*(x+2)) }) + 
  stat_function(data = data_frame(x=x, color = factor(2)),
                fun = function(x) { -3 - sqrt(-8*(x+2)) }) + 
  scale_color_manual(labels = lbs,
                     name = "",
                     values = c(scales::hue_pal()(2)[1], 
                                scales::hue_pal()(2)[2])) + 
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0)
g
```